---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# erahumed

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![R-CMD-check](https://github.com/erahumed/erahumed/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/erahumed/erahumed/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/erahumed/erahumed/graph/badge.svg?token=72POLBUEUR)](https://codecov.io/gh/erahumed/erahumed)
<!-- badges: end -->

The `{erahumed}` R package provides the infrastructure for the 
[ERAHUMED Decision Support System](https://www.erahumed.com/decision-support-system/)
 (DSS), that consists of:

* **Data**. The inputs for the DSS are packaged as R `data.frame`s imported 
with `{erahumed}`.
* **Analysis**. The entire modeling pipeline is implemented by R functions.
* **Visualization**. All model outputs are R objects with methods for data 
exploration and visualization.
* **Interactive Interface**. All of the above has a graphical interface 
implemented as a Shiny application, which is also bundled with the package.

## Installation

Simple installation instructions are not available yet. When the present 
repository goes public, you will be able to install `{erahumed}` using 
`devtools::install_github()`. For the moment, you can either:

* Clone this repository, open the RStudio package project and use 
`devtools::install()`.

* Contact the maintainers.

## Usage

### Example 1: Using the interactive dashboard

The graphical interface to the ERAHUMED DSS can be accessed through the 
following R command:

```{r eval=FALSE, include=TRUE}
erahumed::launch_app()
```

This will open the DSS dashboard in your default browser, from where you can 
explore model outputs in a user friendly manner.

### Example 2: command line interface to the models

TODO [#81](https://github.com/erahumed/erahumed/issues/81).

## Getting help

If you have issues running `{erahumed}` or want to suggest an improvement, 
please [file an issue on Github](https://github.com/erahumed/erahumed/issues).
