---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# erahumed

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![R-CMD-check](https://github.com/erahumed/erahumed/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/erahumed/erahumed/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/erahumed/erahumed/graph/badge.svg?token=72POLBUEUR)](https://codecov.io/gh/erahumed/erahumed)
<!-- badges: end -->

The `{erahumed}` R package provides the infrastructure for the 
[ERAHUMED Decision Support System](https://www.erahumed.com/decision-support-system/)
 (DSS), that consists of:

* **Data**. The inputs for the DSS are packaged as R `data.frame`s imported 
with `{erahumed}`.
* **Analysis**. The entire simulation pipeline is implemented by R functions.
* **Visualization**. All simulation outputs are R objects with methods for data 
exploration and visualization.
* **Interactive Interface**. All of the above has a graphical interface 
implemented as a Shiny application, which is also bundled with the package.

## Installation

You can install the latest release of `{erahumed}` from Github, by running the following command in your local R session:

```{r eval=FALSE, include=TRUE}
remotes::install_github("erahumed/erahumed@*release")
```

If you want the development version (notice: this may be unstable), simply omit the `@*release` reference, as in: 

```{r eval=FALSE, include=TRUE}
remotes::install_github("erahumed/erahumed")  # development version
```

## Usage

### Example 1: Using the interactive dashboard

The graphical interface to the ERAHUMED DSS can be accessed through the 
following R command:

```{r eval=FALSE, include=TRUE}
erahumed::launch_app()
```

This will open the DSS dashboard in your default browser, from where you can 
explore simulation outputs in a user friendly manner.

### Example 2: command line interface to simulations

```{r}
library(erahumed)
```


The following example illustrates the workflow for manually running the ERAHUMED
simulation chain, and extracting the outputs of the various simulation layers. For further information, see the 
[main package vignette](https://erahumed.github.io/erahumed/articles/erahumed-workflow.html).

The pipeline always starts by initializing an ERAHUMED simulation, via:

```{r}
simulation <- erahumed_simulation()
simulation
```

This is the main abstraction that `{erahumed}` uses to collect the various 
layers involved in the simulation chain of the DSS. As the output above shows, 
its layers are initialized but not computed yet. 

We can modify the configuration of specific layers through the `setup_*()` 
functions, for instance:

```{r}
simulation <- simulation |>
  setup_hbp(ideal_flow_rate_cm = 2.5) |>
  setup_ct(dact_m = 0.2)
```

In order to actually compute the layers, we use `run_simulation()`:

```{r}
simulation <- simulation |>
  run_simulation(layer = "hba")  # Run simulation until the HBA layer
```

Simulations are composed of the following layers (in dependency order, from 
upstream to downstream):

* INP [`setup_inp()`]: INPut data.
* HBA [`setup_hba()`]: Hydrological Balance of the Albufera lake.
* HBP [`setup_hbp()`]: Hydrological Balance of rice Paddy clusters.
* CA [`setup_ca()`]: Chemical Applications. 
* CT [`setup_ct()`]: Chemical Transport.

In order to inspect the results of a given layer, we use:

```{r}
get_layer(simulation, "hba")
get_layer_parameters(simulation, "hba")
get_layer_output(simulation, "hba") |> head()
```

For more information on the simulation interface, you can consult `?erahumed_simulation_interface`.

## Getting help

If you have issues running `{erahumed}` or want to suggest an improvement, 
please [file an issue on Github](https://github.com/erahumed/erahumed/issues).

An internal documentation, aimed at potential contributors to this project, is available at [Github Wiki](https://github.com/erahumed/erahumed/wiki).
