---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# erahumed

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![R-CMD-check](https://github.com/erahumed/erahumed/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/erahumed/erahumed/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/erahumed/erahumed/graph/badge.svg?token=72POLBUEUR)](https://codecov.io/gh/erahumed/erahumed)
[![Website](https://img.shields.io/badge/Website-here-blue)](https://erahumed.github.io/erahumed/)
<!-- badges: end -->

The `{erahumed}` R package provides the infrastructure for the 
[ERAHUMED Decision Support System](https://www.erahumed.com/decision-support-system/)
 (DSS), that consists of:

* **Data**: The DSS inputs are pre-packaged as R `data.frame`s and seamlessly 
imported with `{erahumed}`.
* **Analysis**. The entire simulation workflow is implemented as a set of R 
functions, enabling flexible and reproducible analysis.
* **Outputs**. Simulation results are returned as well-structured `data.frame`s, 
ready for further exploration and processing in R.
* **Interactive Interface**. A Shiny-based graphical interface, with 
predefined visualizations for model outputs, is bundled with the package, 
allowing users to interactively explore the results.

More information on the ERAHUMED project can be found on the [main project's website](https://www.erahumed.com/). The full documentation of the `{erahumed}` 
R package is hosted at [erahumed.github.io/erahumed](https://erahumed.github.io/erahumed/).

## Installation

You can install the latest release of `{erahumed}` from Github, by running the 
following command in your local R session:

```{r eval=FALSE, include=TRUE}
install.packages("remotes")  # If necessary
remotes::install_github("erahumed/erahumed", ref = remotes::github_release())
```

If you want the development version (notice: this may be unstable), simply omit 
the reference: 

```{r eval=FALSE, include=TRUE}
remotes::install_github("erahumed/erahumed")  # development version
```

## Usage

### Example 1: Using the interactive dashboard

The graphical interface to the ERAHUMED DSS can be accessed through the 
following R command:

```{r eval=FALSE, include=TRUE}
erahumed::launch_dss()
```

This will open the DSS dashboard in your default browser, from where you can 
explore simulation outputs in a user friendly manner.

### Example 2: command line interface to simulations

```{r}
library(erahumed)
```


The following example illustrates the workflow for manually running the ERAHUMED
simulation chain, and extracting the outputs of the various simulation layers. 
For more detailed information, see the 
[main package vignette](https://erahumed.github.io/erahumed/articles/erahumed-workflow.html).

Simulations are run via:

```{r}
simulation <- erahumed_simulation(
  date_start = "2020-01-01",
  date_end = "2020-12-31",
  variety_prop = c(J.Sendra = 0.7, Bomba = 0.15, Clearfield = 0.15)
)
```

where the parameters of the simulation are set through the arguments of 
`erahumed_simulation()` (see `?erahumed_simulation` or [this article](https://erahumed.github.io/erahumed/articles/simulation-inputs.html) 
for a full list of parameters.

Results can be inspected through:

```{r}
get_results(simulation, 
            component = "exposure",  # either "hydrology", "exposure", or "risk"  
            element = "lake"         # either "lake", "ditch", or "cluster"
            ) |>
  head()
```

## Getting help

The full documentation of the `{erahumed}` R package is hosted at [erahumed.github.io/erahumed](https://erahumed.github.io/erahumed/).

If you have issues running `{erahumed}` or want to suggest an improvement, 
please [file an issue on Github](https://github.com/erahumed/erahumed/issues).

An internal documentation, aimed at potential contributors to this project, is available at [Github Wiki](https://github.com/erahumed/erahumed/wiki).
