---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# erahumed

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![R-CMD-check](https://github.com/erahumed/erahumed/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/erahumed/erahumed/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/erahumed/erahumed/graph/badge.svg?token=72POLBUEUR)](https://codecov.io/gh/erahumed/erahumed)
<!-- badges: end -->

The `{erahumed}` R package provides the infrastructure for the 
[ERAHUMED Decision Support System](https://www.erahumed.com/decision-support-system/)
 (DSS), that consists of:

* **Data**. The inputs for the DSS are packaged as R `data.frame`s imported 
with `{erahumed}`.
* **Analysis**. The entire modeling pipeline is implemented by R functions.
* **Visualization**. All model outputs are R objects with methods for data 
exploration and visualization.
* **Interactive Interface**. All of the above has a graphical interface 
implemented as a Shiny application, which is also bundled with the package.

## Installation

You can install the latest release of `{erahumed}` from Github, by running the following command in your local R session:

```{r eval=FALSE, include=TRUE}
remotes::install_github("erahumed/erahumed@*release")
```

If you want the development version (notice: this may be unstable), simply omit the `@*release` reference, as in: 

```{r eval=FALSE, include=TRUE}
remotes::install_github("erahumed/erahumed")  # development version
```

## Usage

### Example 1: Using the interactive dashboard

The graphical interface to the ERAHUMED DSS can be accessed through the 
following R command:

```{r eval=FALSE, include=TRUE}
erahumed::launch_app()
```

This will open the DSS dashboard in your default browser, from where you can 
explore model outputs in a user friendly manner.

### Example 2: command line interface to the models

```{r}
library(erahumed)
```


The following example illustrates the workflow for manually running the ERAHUMED
model chain, and extracting the outputs of the various model layers.

The pipeline always starts by initializing a new ERAHUMED model, via:

```{r}
model <- erahumed_simulation()
model
```

This is the main abstraction that `{erahumed}` uses to collect the various 
layers involved in the modeling chain of the DSS, and is initially blank.

Next, we populate model layers, which is achieved through the set of
`compute_*()` functions:

* `compute_inp()` - INPut data.
* `compute_hba()` - Hydrological Balance of the Albufera lake.
* `compute_hbp()` - Hydrological Balance of rice Paddy clusters.
* `compute_ca()` - Chemical Applications. 
* `compute_ct()` - Chemical Transport.

For instance, we can compute the first two layers with default arguments (*i.e.* 
default input data and parameters for hydrological balance computation) as 
follows:

```{r}
model <- model |> compute_inp() |> compute_hba()
model
```

We can extract and inspect model layers as illustrated in the following 
examples:

```{r}
hba(model)  # Extract the HBA model layer
hba(model) |> 
  layer_output() |>  # Get output data.frame
  tibble::as_tibble()    # Just for pretty printing
```

layers have plot methods that can help data exploration:

```{r eval=FALSE, include=TRUE}
plot(hba(model), variable = "outflow_total")
```

(The output is an interactive plot, not shown here. Try yourself!)

Extracting a layer that has not been computed yet yields a `NULL` result, 
*e.g.*:

```{r}
hbp(model)
```

If we attempt to compute a downstream layer whose dependencies have not been
defined yet, we get an informative error. For example:

```{r, error=TRUE}
compute_ca(model)
```

Also, computing a layer erases the output of its downstream dependencies (if
any):

```{r}
compute_inp(model) |> hba()
```

The order of model layers in the list above is the logical one. Each layer depends on the previous ones (referred to as "upstream"), and is a dependency of the subsequent ones (referred to as "downstream").

```{r}
model <- compute_inp(model)
model
```


## Getting help

If you have issues running `{erahumed}` or want to suggest an improvement, 
please [file an issue on Github](https://github.com/erahumed/erahumed/issues).

An internal documentation, aimed at potential contributors to this project, is available at [Github Wiki](https://github.com/erahumed/erahumed/wiki).
