---
title: "Simulation workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{erahumed-workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This document illustrates the simulation workflow of ERAHUMED. Specifically, we 
will cover:

* How to setup and run a simulation.
* How to extract and analyze simulation results.

This guide is addressed to users working with the command-line (*i.e.* R) 
interface of `{erahumed}`.

```{r load-erahumed}
library(erahumed)
```

## Bird's eye view

Running a ERAHUMED simulation involves several layers of computation, which we 
formalize through the concept of simulation *layers*. Concretely, simulations
are composed of the following layers:

* INP: INPut data.
* hbl: Hydrological Balance of the Albufera lake.
* hbc: Hydrological Balance of rice Paddy clusters.
* CA: Chemical Applications. 
* CT: Chemical Transport.

Each layer depends on the previous ones and is a dependency of the subsequent
ones (in the order above). The goal of the following sections is to teach you
how to operate with simulations and layers in `{erahumed}`.

## The simulation workflow

We can create a new simulation with:

```{r}
s <- erahumed_simulation()
s
```

The object `s` is, behind the scenes, an R list that contains simulation layers.
The list of ERAHUMED layers can be obtained through:

```{r}
erahumed_layers()
```

We can extract individual layers as follows:

```{r}
get_layer(s, "hbl")  # Extract the hbl layer
```

As revealed by the outputs above, the layers of the simulation are not yet 
computed at this stage. This is so because the only result of 
`erahumed_simulation()` is to *initialize* a simulation, using default values
for the parameters required by the various simulation layers. We show in the
following how to setup individual layers, run the simulation and extract 
results.

### Setup

Setup of individual layers is achieved through the family of `setup_*()` 
functions, that receive a simulation as input, and return a simulation with the 
updated setup. 

For instance, to configure the height threshold parameter used in hbc layer 
computations (see `?setup_hbc`), we would use:

```{r}
s <- setup_hbc(s, height_thresh_cm = 1) 
```

Similarly, to limit the date range of the simulation, we can provide a filtered  
`outflows_df` data-set to the INP layer:

```{r}
outflows_df <- albufera_outflows |>
  (\(.) .["2020-01-01" <= .$date & .$date <= "2020-12-31", ])()
s <- setup_inp(s, outflows_df = outflows_df)
```

The available parameters for each layer are the arguments of the corresponding
`setup_*()` function, and the default values set by `erahumed_simulation()` 
correspond to the defaults of the `setup_*()` arguments. The current layer 
parameters of a simulation can be obtained through:

```{r}
get_layer_parameters(s, "ct")  # Get parameters of CT layer
```

### Run

Once we are ready with our simulation setup, we can compute the various layers,
using the `run_simulation()` function. In order to speed up iterations,  
`{erahumed}` gives you the option to compute just a specific layer with all its 
upstream dependencies. For example:

```{r}
s <- run_simulation(s, layer = "hbc")
s
```

would compute only the INP, hbl and hbc layers. If you simply want to compute 
all layers, you can just call:

```{r}
s <- run_simulation(s)
s
```
(notice that this will not recompute the INP, hbl and hbc layers, since these 
results are already stored in the input simulation object).

### Analyse

Once we have setup and ran a simulation, we want to analyse its result, which 
will be essentially time-series data-sets (possibly with a spatial component
identified by a `cluster_id`).

You can extract the output (a `data.frame`) of a given layer as follows:

```{r}
hbc_df <- get_layer_output(s, "hbc")  # Get output data.frame of hbc layer
head(hbc_df)
```

From here on, the analysis may proceed in the way you find more convenient. 
For instance, in the chunk below I create a plot of water levels for a set of 
clusters with similar features, using `dplyr` and `ggplot2`:

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

ditch <- "d4"
tancat <- FALSE
variety <- "Clearfield"

clusters_df <- hbc_df |>
  filter(ditch == !!ditch, tancat == !!tancat, variety == !!variety)

avg_df <- clusters_df |>
  group_by(date) |>
  summarise(height_eod_cm = mean(height_eod_cm))

ggplot() +
  geom_line(
    data = clusters_df,
    mapping = aes(x = date, y = height_eod_cm, group = cluster_id),
    color = "black", linewidth = 0.1, alpha = 0.2) +
  geom_line(
    data = avg_df, 
    mapping = aes(x = date, y = height_eod_cm),
    color = "black"
    ) +
  xlab("Date") + ylab("Height [cm]") + 
  ggtitle("Cluster simulated water levels",
          paste("Ditch:", ditch, "- Tancat:", tancat, "- Variety:", variety)
          )
```

## Predefined plotting methods

`{erahumed}` comes equipped with a few basic plotting methods that return 
interactive visualizations of the various model layers. These can be used by 
accessing the relevant model layer through `get_layer()`, for example:

```{r}
get_layer(s, "hbl") |>
  plot(variable = "volume")

```

The documentation of the various plotting methods can be found under 
`?plot.erahumed_<layer_name>`, *e.g.* `?plot.erahumed_hbl`.

## Further information

Further details will appear in this and possibly other vignettes. For 
specific problems, you can [file an issue on Github](https://github.com/erahumed/erahumed/issues).
