---
title: "Simulation workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{erahumed-workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document illustrates the simulation workflow of ERAHUMED. Specifically, we 
will cover:

* How to setup and run a simulation.
* How to extract and analyze simulation results.

This guide is addressed to users working with the command-line (*i.e.* R) 
interface of `{erahumed}`.

```{r load-erahumed}
library(erahumed)
```

ERAHUMED simulation can be run with:

```{r}
sim <- erahumed_simulation()
sim
```
The command above runs a simulation with the default model parameters. These can be customized through the arguments of `erahumed_simulation()`, for instance:

```{r}
sim2 <- erahumed_simulation(
  variety_prop = c(J.Sendra = 0.6, Bomba = 0.2, Clearfield = 0.2),
  ideal_flow_rate_cm = 10,
  seed = 841
)
```

The full set of simulation parameters is documented in `?erahumed_simulation` (see [here](https://erahumed.github.io/erahumed/articles/simulation-inputs.html) for
a table format).

Once we are ready with our simulation setup, in order to actually run the 
simulation we use:

Simulation results are extracted as follows:

```{r}
lake_hydrology_df <- get_results(sim, component = "hydrology", element = "lake")
cluster_hydrology_df <- get_results(sim, component = "hydrology", element = "cluster")
cluster_exposure_df <- get_results(sim, component = "exposure", element = "cluster")
```
These are provided in the form of `data.frame`s, for instance:

```{r}
head(cluster_hydrology_df)
```

From here on, the analysis may proceed in the way you find more convenient. 
For instance, in the chunk below I create a plot of water levels for a set of 
clusters with similar features, using `dplyr` and `ggplot2`:

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

ditch <- "d4"
tancat <- FALSE
variety <- "Clearfield"

clusters_df <- cluster_hydrology_df |>
  filter(ditch == !!ditch, tancat == !!tancat, variety == !!variety)

avg_df <- clusters_df |>
  group_by(date) |>
  summarise(height_eod_cm = mean(height_eod_cm))

ggplot() +
  geom_line(
    data = clusters_df,
    mapping = aes(x = date, y = height_eod_cm, group = cluster_id),
    color = "black", linewidth = 0.1, alpha = 0.2) +
  geom_line(
    data = avg_df, 
    mapping = aes(x = date, y = height_eod_cm),
    color = "black"
    ) +
  xlab("Date") + ylab("Height [cm]") + 
  ggtitle("Cluster simulated water levels",
          paste("Ditch:", ditch, "- Tancat:", tancat, "- Variety:", variety)
          )
```

## Further information

Further details will appear in this and possibly other vignettes. For 
specific problems, you can [file an issue on Github](https://github.com/erahumed/erahumed/issues).
