---
title: "Simulation workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{erahumed-workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document illustrates the simulation workflow of ERAHUMED. Specifically, we 
will cover:

* How to setup and run a simulation.
* How to extract and analyze simulation results.

This guide is addressed to users working with the command-line (*i.e.* R) 
interface of `{erahumed}`.

```{r load-erahumed}
library(erahumed)
```

We initialize a new ERAHUMED simulation with:

```{r}
sim <- erahumed_simulation()
sim
```
At this stage, `sim` is simply a list of simulation parameters, all initialized 
to their default values. To customize these parameters, we use 
`setup_hydrology()`, `setup_exposure()`, and `setup_risk()`. These functions 
provide access to the specific subset of parameters associated with 
the hydrology, exposure, or risk components of the simulation, respectively.
For instance, the following command modifies the "height threshold" numeric 
parameter and provides a custom `outflows_df` input data-set with a reduced 
date range (see `?setup_hydrology` for the detailed documentation of these 
parameters):

```{r}
outflows_df <- albufera_outflows |>
  (\(.) .["2020-01-01" <= .$date & .$date <= "2020-12-31", ])()
sim <- sim |> 
  setup_hydrology(height_thresh_cm = 1, outflows_df = outflows_df)
```

Once we are ready with our simulation setup, in order to actually run the 
simulation we use:

```{r}
sim <- sim |> 
  run_simulation()
sim
```
In order to extract simulation results we use:

```{r}
lake_hydrology_df <- get_results(sim, component = "hydrology", element = "lake")
cluster_hydrology_df <- get_results(sim, component = "hydrology", element = "cluster")
cluster_exposure_df <- get_results(sim, component = "exposure", element = "cluster")
```
These are provided in the form of `data.frame`s, for instance:

```{r}
head(cluster_hydrology_df)
```

From here on, the analysis may proceed in the way you find more convenient. 
For instance, in the chunk below I create a plot of water levels for a set of 
clusters with similar features, using `dplyr` and `ggplot2`:

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

ditch <- "d4"
tancat <- FALSE
variety <- "Clearfield"

clusters_df <- cluster_hydrology_df |>
  filter(ditch == !!ditch, tancat == !!tancat, variety == !!variety)

avg_df <- clusters_df |>
  group_by(date) |>
  summarise(height_eod_cm = mean(height_eod_cm))

ggplot() +
  geom_line(
    data = clusters_df,
    mapping = aes(x = date, y = height_eod_cm, group = cluster_id),
    color = "black", linewidth = 0.1, alpha = 0.2) +
  geom_line(
    data = avg_df, 
    mapping = aes(x = date, y = height_eod_cm),
    color = "black"
    ) +
  xlab("Date") + ylab("Height [cm]") + 
  ggtitle("Cluster simulated water levels",
          paste("Ditch:", ditch, "- Tancat:", tancat, "- Variety:", variety)
          )
```

## Further information

Further details will appear in this and possibly other vignettes. For 
specific problems, you can [file an issue on Github](https://github.com/erahumed/erahumed/issues).
